---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# codecat

<!-- badges: start -->
<!-- badges: end -->

The goal of codecat is to provide a minimal toolkit for defining and using codebooks (also called data dictionaries) which specify factor levels and labels for one or more (usually categorical) variables in a data set.

## Installation

You can install the development version of codecat like so:

``` r
remotes::install_github("djnavarro/codecat")
```

## Example

```{r load-package}
library(codecat)
library(tibble)
```

To specify a coding scheme for a variable, use the `code()` function. A collection of coding schemes that should be applied to a data set is called a **codebook**, and is created using `codebook()`:

```{r define-codebook}
codebook(
  code("SEX", levels = c(0, 1), labels = c("Male", "Female")),
  code("STUDY", levels = c(1, 2, 3), labels = c("S1", "S2", "S3"))
)
```

Apply it to data using `encode()`:

```{r use-codebook}
dat <- tibble(
  STUDY = c(1, 1, 1, 2, 2, 2),
  SEX   = c(0, 0, 1, 1, 1, 0),
  AGE   = c(32, 18, 64, 52, 26, 80)
)

labels <- codebook(
  code("SEX", levels = c(0, 1), labels = c("Male", "Female")),
  code("STUDY", levels = c(1, 2, 3), labels = c("S1", "S2", "S3"))
)

encode(dat, labels)
```

To restore original levels, use `decode()`:

```{r decode-data}
dat2 <- encode(dat, labels)
decode(dat2, labels)
```
